---
title: "Regression models for updated results"
author: "Nathan Hahn"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(tidyverse)
library(MuMIn)
library(ggeffects)
library(DescTools)
```



## 1. Assess impact of age, sex, and home range on ag usage

Import individual-year data 
```{r}
# mod.df is the change.df from Tactic_Change_Models.R
mod.df <- read.csv("./GMM/results/individual_year_classification.csv")

output <- readRDS('./GMM/movdata/GMEcollars_003_usedClust_2020-10-30.rds')

#knitr::kable(head(mod.df))
```

Prep data

- Hypothesize that there should be a strong positive association between crop usage and male elephants, and crop usage and old adults. Refernce levels set for sex and age class to look for this relationship 
```{r}
# normalize mcp
# mod.df$year.mcp.area <- (mod.df$year.mcp.area - min(mod.df$year.mcp.area))/
#   (max(mod.df$year.mcp.area) - min(mod.df$year.mcp.area))


#mod.df$year.mcp.area <- log(mod.df$year.mcp.area)

mod.df$subject_sex <- as.factor(mod.df$subject_sex) # females as reference level

mod.df$subject_ageClass <- as.factor(mod.df$subject_ageClass) 
mod.df$subject_ageClass <- relevel(mod.df$subject_ageClass, ref = "young adult") # young adults as reference level


# calculate mean daily displacement
## Mean and stdev of cummulative daily distance moved (displacement)
# get daily aggregate stats on movement
output$ymd <- as.Date(output$date, "%Y/%m/%d")
stat <- function(x) c(sum = sum(x))
ag.daily <- as.data.frame(aggregate(dist ~ ymd + subject_name, output, stat)) # aggregated stats by day

# bind daily displacment to full dataset
displacement <- inner_join(output, ag.daily, by = c("ymd", "subject_name")) %>%
  rename(daily.disp = dist.y) %>%
  mutate(tactic.season = as.numeric(tactic.season), year.cuts = as.numeric(year.cuts)) %>%
  group_by(subject_name, year.cuts, tactic.season) %>%
  summarise(n = n(),
            mean = mean(daily.disp, na.rm = TRUE),
            sd = sd(daily.disp, na.rm = TRUE))

dim(displacement)

mod.df <- inner_join(mod.df, displacement, by = c("subject_name", "year.cuts", "tactic.season"))
mod.df <- rename(mod.df, mu.daily.disp = mean)

## Quick summary stats for each covariate

# summary of mcp by tactic class as a proxy for how home range should increase with increasing ag usage
aggregate(year.mcp.area ~ tactic.season, data = mod.df, FUN = 'mean')

# summary stats for sex - shows slight difference
aggregate(year.mean ~ subject_sex, data = mod.df, FUN = 'mean')

# summary stats for age class - shows very little difference in use 
aggregate(year.mean ~ subject_ageClass, data = mod.df, FUN = 'mean')


```

Fit models

- Models are fit to individual-years, so we use a mixed effects model with random intercept by individual to account for repeated measures on the individual. M4 (sex and home range size) produces the best model fit, and drops age class. 

- In this model, increased home range is associated with increased ag usage; b_hr = `0.042`. Males use more agriculture than females but the difference is not as large as would be expected based on previous research; b_male = `0.026 + 0.035`, b_female = `0.035`
```{r}
m1 <- lmer(year.mean ~ subject_sex + subject_ageClass + log(year.mcp.area) + (1|subject_name), data = mod.df, REML = FALSE)
#summary(m1)

m2 <- lmer(year.mean ~ subject_sex + subject_ageClass + (1|subject_name), data = mod.df, REML = FALSE)
#summary(m2)

m3 <- lmer(year.mean ~ subject_sex + log(year.mcp.area) + (1|subject_name), data = mod.df, REML = FALSE)
summary(m3)

m3.disp <- lmer(year.mean ~ subject_sex + log(year.mcp.area) + log(mu.daily.disp) + (1|subject_name), data = mod.df, REML = FALSE)
summary(m3.disp)

m4 <- lmer(year.mean ~ subject_sex + (1|subject_name), data = mod.df, REML = FALSE)
#summary(m4)

m5 <- lmer(year.mean ~ subject_ageClass + (1|subject_name), data = mod.df, REML = FALSE)
#summary(m5)

m6 <- lmer(year.mean ~ (1|subject_name), data = mod.df, REML = FALSE)

knitr::kable(AICc(m1, m2, m3, m4, m5, m6))

sjPlot::tab_model(m3)
```

Looking at residual plots, the model assumptions may be iffy. Residuals are clustered tightly at low levels of ag use, and then quickly disperse. 

This also contributed to overfitting of the multivariate cluster models. I guess this is an inherent issue with trying to model outliers (habitual crop raiders). Including tactic in the model exacerbates the heteroscedasticity. 

A plot of fitted and actual mean ag use values looks good though. 

```{r}
# check residuals
plot(m3)

# box-plots of residuals by Subject
plot(m3, subject_name ~ resid(., scaled=TRUE))

# checking fitted with mean looks good
plot(m3, year.mean ~ fitted(.), abline = c(0,1))
```

Use the fitted model, see that males use more ag than females, but not significantly more.
```{r}
dat <- ggpredict(m3.disp, terms = c("year.mcp.area"))
plot(dat)

```

## 2. Activity overlap across tactics: logistic regression
Differentiation in risk tolerance and the deviation in state activity budgets between tactics is expected to translate into differentiation in circadian activity patterns during ag and non-ag movement phases. We constructed state-level activity pattern densities for ag and non-ag movement phases, calculated coefficients of overlap of the density distributions across tactics, and then compare coefficients for ag and non-ag phases. We expect there to be a low overlap in activity between tactics in agriculture, because of the differentiation in risk tolerance between the tactics. 

Load output from coefficient of overlap tests and visualize group-level means and 95% confidence intervals.
```{r}
comps.output <- read.csv('overlap_comps_output.csv') %>% # no longer including comps using all data (baseline)
  filter(AgWindow != "Baseline") %>% droplevels()

# prep data
comps.output$AgWindow <- as.factor(comps.output$AgWindow)

knitr::kable(head(comps.output))

## Visualize overlap comp results

# get tactic-level means and CIs
comp.sum <- comps.output %>%
  group_by(AgWindow, state.names) %>%
  mutate(state.names = factor(state.names, levels = c("Encamped", "R-A Search", "Exploratory")),
         AgWindow = factor(AgWindow, levels = c( "Non-Ag Window", "Ag Window"))) %>%
  summarise(n = n(),
            grand.mean.OV = mean(estOV),
            lwr.ci = MeanCI(estOV, method = "boot", type = "norm", R=1000)[2],
            upr.ci = MeanCI(estOV, method = "boot", type = "norm", R=1000)[3])

knitr::kable(comp.sum)

# plot comps
ggplot(data = comp.sum, aes(x = state.names, y = grand.mean.OV)) +
  geom_pointrange(aes(ymin = lwr.ci, ymax = upr.ci, color = AgWindow), position = position_dodge(0.4)) + scale_color_manual(values = c("#F8766D", "#00BFC4"))
  
```

Fit model
- Logistic model fit using overlap estimates and states, with Ag Window (0 = non-ag, 1 = ag) as the response

- Tested an interaction between overlap and state, given that the differences between Ag Window and Non-Ag movement phases, because mean overlaps are not all in the same direction. 

- Logistic regression shows that overlap of activity patterns does not change significantly by ag window, suggesting that elephant circadian patterns are similar when inside and outside of agriculture, irrespective of tactic. Reflects similar risk perceptions across elephants, desipite large differences in risk tolerance. 

- Note that M1 (no inclusion of state) has the lowest AICc, but prevents looking at differences by state
```{r}

# create model dataframe, excluding comparisons using All Data
mod.df <- comps.output

mod.df$state <- as.factor(mod.df$state.names) 
mod.df$state <- factor(mod.df$state, levels = c("Encamped", "R-A Search", "Exploratory"))
mod.df$AgWindow <- as.factor(mod.df$AgWindow)
mod.df$AgWindow <- relevel(mod.df$AgWindow, ref = "Non-Ag Window") # set non-ag window as 0 for interpretation
mod.df$comp <- as.factor(mod.df$comp)


# m1 <- glm(AgWindow ~ estOV, data = mod.df, family = binomial(link = 'logit'))
# #summary(m1)
# 
# # state added with reference level for state 1
# m2 <- glm(AgWindow ~ estOV + state, data = mod.df, family = binomial(link = 'logit'))
# #summary(m2)
# 
# # state added as an interaction term
# m3 <- glm(AgWindow ~ estOV*state, data = mod.df, family = binomial(link = 'logit'))
# 
# knitr::kable(AICc(m1, m2, m3))
```

m2 shows that overlap is significantly higher among tactics in foraging state, and significantly lower in encamped state. 
Tactic comparison estimates overlapped zero, indicating no differences in activity patterns between tactics when inside and outside of ag. 

```{r}

mod.df$state <- relevel(mod.df$state, ref = "Encamped")
mod.df$comp <- relevel(mod.df$comp, ref = "1-4")

#m1 <- glm(estOV ~ state, data = mod.df)
m2 <- glm(estOV ~ AgWindow + state + comp, data = mod.df)
#m3 <- lmer(estOV ~ state + comp + (1|response), data = mod.df, REML = FALSE) # singular fit

summary(m2)

# differences in overlap between states. highest in S2 - Foraging
anova(m2)

drop1(m2, test='Chisq')

library(lmtest)
library(sjPlot)

# plot
dat <- ggpredict(m2, c("state"))
plot(dat)

# table
tab_model(m2, auto.label = TRUE)
```


## 3. Activity pattern changes within tactics based on agricultural usage

Results from the logistic regression showed that individuals in each tactic have similar activity patterns when in and outside of agriculture. Given this similarity, we looked at within-tactic differentiation in activity patterns during agricultural and non-agricultural use periods. 

- The most change in activity pattern (defined as lowest overlap coefficient) comes in the encamped state. 

- Unsure of what further statistical test to do here. The table points at the low within-tactic activity overlap between ag and non-ag movement phases. See correlation test below comparing change in activity budget pcts with the overlap estimates. 

```{r}
#comps.output <- read.csv('within_tactic_comps.csv') 
comps.output <- read.csv('indv.ag.comp.csv')


# prep data
comps.output$state <- as.factor(comps.output$state)
comps.output$state <- relevel(comps.output$state, ref = "1") # RA search as ref level
contrasts(comps.output$state)

comps.output$tactic <- as.factor(comps.output$tactic) 

## Visualize overlap comp results

# get tactic-level means and CIs
# comp.table <- comps.output %>%
#   select(state, tactic, est.OV = estOV, lwr.ci = lwr, upr.ci = upr)
# knitr::kable(comp.table)

indv.comp.table <- comps.output %>% group_by(tactic, state) %>%
  summarise(mean = mean(estOV),
            sd = sd(estOV))
knitr::kable(indv.comp.table)

# plot comps
# ggplot(data = comp.table, aes(x = factor(state), y = est.OV)) +
#   geom_pointrange(aes(ymin = lwr.ci, ymax = upr.ci, color = factor(tactic)), position = position_dodge(0.4)) +
#   xlab("Sate") + ylab("Est. Overlap between Ag and Non-Ag activity pattern")
```

Model fitting - individual

Fitting a mixed effects regression of the within-individual overlap between ag and non-ag movement phases, we find that the overlap differs significantly between states (lowest in encamped, and significantly higher in foraging and exploratory). The overlap between ag and non-ag phases are similar across all tactics. Combined with the model finding no differences in activity overlap across tactics when inside and outside agriculture, this suggests similar activity and similar activity shifts by all tactics in response to agriculture. Reflects similarity in percieved risk. 

Lower overlap in encamped could be occuring because resting patterns are flipped to avoid humans. 

```{r}
## Want to make inference on tactic and state, so we just fit the model to both rather than using model selection to drop tactic

#m0 <- lmer(estOV ~ (1|subject_name), data = comps.output, REML = FALSE)
m1 <- lmer(estOV ~ tactic + state + (1|subject_name), data = comps.output, REML = FALSE)
#m2 <- lmer(estOV ~ state + (1|subject_name), data = comps.output, REML = FALSE)

#AIC(m0,m1,m2)

summary(m1)

confint(m1)

dat <- ggpredict(m1, terms = c("state"))
plot(dat)

library(MuMIn)
r.squaredGLMM(m1)

library(lmerTest)
drop1(m1,test="Chisq")
```

Estimates of overlap and activity budgets within tactics show that individuals change their activity based on ag vs. non-ag movement phases. Because these metrics are produced using the same comparison framework, we can look at how the changes in these two measures are related. Change in activity budgets are calculated as the absolute difference in the percent time spent in a given state, within the tactic.

We would expect that as overlap between ag and non-ag activity pattern increases (becomes more similar), the absolute difference in ag and non-ag activity budgets would decrease.  

The graph and correlation tests below show that for all states, these metrics are not correlated. The timing and allocation of state-level activity is not linked across the agricultrual gradient. Highlights importance of examining circadian patterns in addition to activity budgets when assessing risk perceptions in heterogenous landscapes. 

```{r}
# load data on activity budget allocations
budget <- read.csv('activity_budget_allocations.csv')

# Prep data
budget$state <- as.numeric(budget$state)
budget$tactic <- as.numeric(factor(budget$tactic, levels = c('Rare', 'Sporadic', 'Seasonal', 'Habitual')))
t <- spread(budget, key = ag.window, value = pct) %>%
  arrange(state, tactic) # order to match comp.table

# Get absolute differences between movement phases by tactic and state
t$abs.diff <- abs(t$`0` - t$`1`)

# combine
comp.table$abs.diff.budget <- t$abs.diff
knitr::kable(comp.table)

# relationship between overlap and budget difference by state. colors are tactics
ggplot(comp.table, aes(x = abs.diff.budget, y = est.OV)) + geom_point(aes(color = tactic)) + facet_wrap(. ~ state) + 
  xlab("abs. difference in budget (pct)") + ylab("estimated overlap (pct)")

# correlation tests by state
split <- split(comp.table, comp.table$state)
cor.test(split[[1]]$est.OV, split[[1]]$abs.diff.budget)
cor.test(split[[2]]$est.OV, split[[2]]$abs.diff.budget)
cor.test(split[[3]]$est.OV, split[[3]]$abs.diff.budget)

```







